% $Id: part1.tex $ #
\documentclass[10pt,titlepage]{book}
\usepackage{makeidx}
\usepackage{graphicx}
\usepackage{booktabs}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{unicode-math}
\usepackage[unicode]{hyperref}
\usepackage{endnotes}

\usepackage{paralist}
\usepackage{relsize}
\usepackage{verbatim}
\usepackage{enumerate}
\usepackage{longtable}
\usepackage{url}

\usepackage{fontspec}

% Greek Characters

\setmainfont{TeX Gyre Pagella}[Ligatures=TeX]
\setmathfont{TeX Gyre Pagella Math}[Ligatures=TeX]
%\setmainfont{TeX Gyre Pagella}
\newfontfamily\greekfont{GFS Didot}

\usepackage[paperwidth=5.25in,paperheight=8in,hmargin={0.75in,0.5in},vmargin={0.5in,0.5in},includehead,includefoot]{geometry}
\hypersetup{pdfauthor={Roger Bishop Jones}}
\hypersetup{pdftitle={An Epistemological Synthesis}}
\hypersetup{colorlinks=true, urlcolor=red, citecolor=blue, filecolor=blue, linkcolor=blue}
%\usepackage{html}
\renewcommand{\thefootnote}{\roman{footnote}}

\newcommand{\ignore}[1]{}
\newcommand{\rbjdeff}[1]{\textbf{\itshape{#1}}}
\newcommand{\rbjdef}[1]{\rbjdeff{#1}\label{LABEL#1}\index{#1}}
\newcommand{\rbjuse}[1]{\hyperref[LABEL#1]{\itshape{#1}}}
\newcommand{\hreg}[2]{\href{#1}{#2}\footnote{\url{#1}}}
\newcommand{\groknote}[1]{\vfil
  {\bf Note for Grok:}
    \begin{quote}
      {\it #1}
    \end{quote}
    \vfil
  }

\makeatletter
\renewcommand\enoteheading{\section*{\notesname}%
    \@mkboth{\notesname}{\notesname}}

\newcommand{\resetnotes}{%
  \ifnum\value{endnote}>0%
  \theendnotes%
  \setcounter{endnote}{0}%
  \else%
  \fi%
}

\newcommand{\chapnote}[1]{%
  \chapter{#1}
  \input{{#1}.tex}
  \resetnotes
}

\makeatother

\pagestyle{plain}
\makeindex

\title{\bf\LARGE An Epistemological Synthesis}
\author{Roger~Bishop~Jones}
\date{\small 2025-03-06}


\begin{document}
%\frontmatter

%\begin{abstract}
% Presentation of the most fundamental aspects of a positivist philosophical system articulated around a position on knowledge representation.
%
%\end{abstract}
                               
\begin{titlepage}
\maketitle

%\vfill

%\begin{centering}

%{\footnotesize
%copyright\ Roger~Bishop~Jones;
%}%footnotesize

%\end{centering}

\end{titlepage}

\ \

\ignore{
\begin{centering}
{}
\end{centering}
}%ignore

\setcounter{tocdepth}{2}
{\parskip-0pt\tableofcontents}

%\mainmatter

\chapter{Introduction}

\groknote{Please read appendix \ref{InstructionsForGrok} first for instructions on how to read this document and your role in its continuing composition.}

This monograph is intended primarily as an articulation of a foundational account of an epistemologically rooted philosophical system.
It is however, practically motivated.
The fundamental ideas presented here are offered, not only as the kernel of a philosophy, but also as a contribution to the most abstract stages in the design of a widely distributed shared repository of knowledge.
A repository which may underpin the advancement of humanity and our progeny (biological or synthetic) into the future.


Synthetic Epistemology\footnote{As that term is used in this monograph.}, and this epistemological synthesis, offers (but does not prescribe) a conception of knowledge for the future.
It is therefore influenced by the authors perception of what that future will be.
This becomes more speculative to the extent that the future is likely to be very different to the present.

Uncertainty about the future is well warranted, for there are many signs that we are at a point of radical inflection in many historical trends.

That evolution began on planet Earth some 4 billion years ago, and over the  next few billion years may ultimately enable the spread intelligent systems across the Milky Way galaxy and into the Cosmos beyond.
The place in which we find ourselves in this ongoing evolution of intelligence marks a major transition in the manner in which intelligence evolves and proliferates.

Two previous similarly momentous moments were \emph{abiogenesis}, the beginnings of life as we know it here on earth, which initiated the biological evolution of DNA coded lifeforms, and the evolution of oral language in \emph{homo sapiens}, when cultural evolution was unleashed and entered on its own hyper-exponential trajectory, enabling the most intelligent species ever seen on earth to dominate and proliferate throughout the ecosphere.

The main features of this special time are the advent of Artificial Intelligence, and its pending application through synthetic biology to the transformation of biological evolution and the accelerated evolution of synthetic intelligent systems.
We are approaching `The Singularity', that moment when the intelligent systems reached through collaborative culural evolution becomes capable itself of continuing evolution by design of increasingly intelligent systems.

A hallmark of such systems, we may anticipate, will be their ability to accumulate knowledge more rapidly, and their ability to exploit that knowledge more effectively, by complex chains of formal reasoning whose predictive power facilitates proliferation.

I am therefore drawn to consider evolution as an epistemic progression in more ways than one.
It is so as an accelerating accumulation of knowledge, and also as the evolution of the cognitive capability which enables that accumulation.

If indeed we advance into that future in which intelligent systems designed by our progeny are spreading across the galaxy, the details of how live has evolved on earth will have diminishing significance and the dominant forms of intelligent system will be those most adept at ``taking root'' in completely alien environments, most able to harvest energy and the material resources to build the computational capability for intelligence and the construction and propulsion of emisaries for the next stage of the proliferation of intelligence.

Not only must the accomplish those remarkable feats, but they must be capable of continuing evolution, on much shorter timescales than the biological evolution which created their human progenitors.
This process will be one of intelligent collaborative advancement of science and technology, building on the fruits of intelligence past held in a shared distributed repository and constributing to the growth of that body of knowledge at the same time as designing the next generation of self-proliferating intelligent systems to move ever more rapidly across the galaxy and beyond.

That respository of knowledge will be indispensable for success in proliferation, and we may therefore expect getting that repository in good shape is something that will happen early in the process, with those unable to exploit and advance its content falling behind in the evolutionary stakes.

It is therefore appropriate to see to what extent it is now possible to anticipate the characteristics of such a repository to secure the competitive advantage that getting there early will bring.

Some of the critical elements are well established in our present culture.
One of those is a philosophical concensus on how scientific knowledge \emph{works} from which we can draw some insights into the characteristics of a suitable repository.
It is no exaggeration that our present prosperity and the possibility that we might move from this planet to other planets, star systems and galaxies, are due to the power of science and technology.
In the establishment of modern scientific methods there have been two major milestones, which shared an emphasis upon the role of deduction in the application of science, while differing perhaps on how that scientific knowledge should be established.

The first of those is Aristotle's Organon and the idea of ``Demonstrative Science'' which he there advances.
The Greek philosophers beginning with Thales in around 600 BC had a broad conception of the scope of philosophy (a love of theoretical knowledge in general), which included mathematics which, as a theoretical discipline they separated from the practice of arithmetical computation and geometrical assay into the business of promulgating abstract theorems obtained by elaborate chains of deductive reasoning from limited and explict starting points.
This method became known as the axiomatic method, and remains (after further refinement) a pillar of modern mathematics.

These methods were very successful in creating a solid body of mathematical knowledge, without controversy and contradition, in stark contrast with the efforts of these pre-socratic philosophers in their attempts to understand the broder cosmos by means of reason rather than appeal to tradition or superstition.
Reason, applied outside the ordely confines of axiomatic mathematics proved incapable of establishing solid results which could survive contradiction by the philosophers who followed.

This contrast between the fruits of reason in mathematics and elsewhere lead made the Greek philosophers to look upon axiomatic mathematics with envy, and to seek successful broadening of the scope of the axiomatic deductive method beyond mathematics.
This we may see in the philosophy of Plato as progressed by his distinguishing two ``worlds'', that of ideal forms, and that of sensory impressions.
It was then held that true knowledge, obtainable by reason, was confined to the world of ideal forms, and that sensory impressions lead no further than opinion.

For Science consists in identifying the fundamental principles in each scientific domain, and proceeding thence by deduction.


\ignore{
The epicenter of this monograph is a conception of finitary declarative language and a relationship of semantic reducibility among these languages, which holds when one language can be translated into another by a computable mapping which preserves abstract semantics.
In the context of these ideas certain 
Because of these characteristics, the general problem of defining the semantics of declarative languages and resoning from knowledge represented in them can be addressed through such reductions.

These quasi-universal declarative languages are candidate representatons for large scale repositories of knowledge.
They are also serve to

\cite{couturat1901logique}

among which certain languages are found to be universal' in the sense that any other language
``quasi-universal'' abstract representation for declarative language and logical truth, with a sound and 'practically complete' formal deductive system.
To this abstract language all other declarative language is reducible

}%ignore

\resetnotes

\chapter{A Philosophical Kernel}\label{PK}

%\input{kernel5.tex}
\resetnotes

\chapter{Evolution}

\cite{donald1991,murray2017evolution}

\ignore{
\chapter{Synthetic Epistemology and Foundational Abstract Semantics}

\input{epistem01.tex}
\resetnotes

\chapter{The Evolution of Declarative Knowledge}\label{EDK}

%\input{edk02.tex}
\resetnotes


\chapter{The Fundamental Triple Trichotomy}\label{FTT}

%\input{ftt01.tex}
\resetnotes
}%ignore

\appendix

\ignore{
\chapter{The Philosophy of Rudolf Carnap}\label{PRC}

\input{carnap01.tex}
\resetnotes

}%ignore

\appendix

\chapter{Instructions for Grok3}\label{InstructionsForGrok}

I have written this monograph with copious assistance from Grok 3, an AI chatbot from xAI.

Notwithstanding the invaluable contribution Grok has made, the writing is all mine.

The following section contains the general instructions which I have used to brief Grok.
Naturally, these instructions have evolved over the course of the writing and this their state at the end.
Different stages in the process demand different instructions, and I have tried to keep instructions of a general character  in a separate file (presented below), with supplementary files giving more specific asks for particular stages.

\section{General Guidance for Grok}

\input{grok3in.tex}

\listoftables
%\listoffigures

%\phantomsection
\addcontentsline{toc}{section}{Bibliography}
\bibliographystyle{rbjfmu}
\bibliography{rbj3}

%\phantomsection
\renewcommand{\indexname}{Index of Defined Terms}
\addcontentsline{toc}{section}{Index of Defined Terms}
{\twocolumn[]
{\small\printindex}}

%\vfill

%\tiny{
%Started 2024-10-19
%}%tiny


\end{document}

% LocalWords:
