% $Id: part1.tex $ #
\documentclass[10pt,titlepage]{book}
\usepackage{makeidx}
\usepackage{graphicx}
\usepackage{booktabs}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{unicode-math}
\usepackage[unicode]{hyperref}
\usepackage{endnotes}

\usepackage{paralist}
\usepackage{relsize}
\usepackage{verbatim}
\usepackage{enumerate}
\usepackage{longtable}
\usepackage{url}

\usepackage{fontspec}

% Greek Characters

\setmainfont{TeX Gyre Pagella}[Ligatures=TeX]
\setmathfont{TeX Gyre Pagella Math}[Ligatures=TeX]
%\setmainfont{TeX Gyre Pagella}
\newfontfamily\greekfont{GFS Didot}

\usepackage[paperwidth=5.25in,paperheight=8in,hmargin={0.75in,0.5in},vmargin={0.5in,0.5in},includehead,includefoot]{geometry}
\hypersetup{pdfauthor={Roger Bishop Jones}}
\hypersetup{pdftitle={An Epistemological Synthesis}}
\hypersetup{colorlinks=true, urlcolor=red, citecolor=blue, filecolor=blue, linkcolor=blue}
%\usepackage{html}
\renewcommand{\thefootnote}{\roman{footnote}}

\newcommand{\ignore}[1]{}
\newcommand{\rbjdeff}[1]{\textbf{\itshape{#1}}}
\newcommand{\rbjdef}[1]{\rbjdeff{#1}\label{LABEL#1}\index{#1}}
\newcommand{\rbjuse}[1]{\hyperref[LABEL#1]{\itshape{#1}}}
\newcommand{\hreg}[2]{\href{#1}{#2}\footnote{\url{#1}}}
\newcommand{\groknote}[1]{\vfil
  {\bf Note for Grok:}
    \begin{quote}
      {\it #1}
    \end{quote}
    \vfil
  }

\makeatletter
\renewcommand\enoteheading{\section*{\notesname}%
    \@mkboth{\notesname}{\notesname}}

\newcommand{\resetnotes}{%
  \ifnum\value{endnote}>0%
  \theendnotes%
  \setcounter{endnote}{0}%
  \else%
  \fi%
}

\newcommand{\chapnote}[1]{%
  \chapter{#1}
  \input{{#1}.tex}
  \resetnotes
}

\makeatother

\pagestyle{plain}
\makeindex

\title{\bf\LARGE An Epistemological Synthesis}
\author{Roger~Bishop~Jones}
\date{\small 2025-03-06}


\begin{document}
%\frontmatter

%\begin{abstract}
% Presentation of the most fundamental aspects of a positivist philosophical system articulated around a position on knowledge representation.
%
%\end{abstract}
                               
\begin{titlepage}
\maketitle

%\vfill

%\begin{centering}

%{\footnotesize
%copyright\ Roger~Bishop~Jones;
%}%footnotesize

%\end{centering}

\end{titlepage}

\ \

\ignore{
\begin{centering}
{}
\end{centering}
}%ignore

\setcounter{tocdepth}{2}
{\parskip-0pt\tableofcontents}

%\mainmatter


\groknote{Please read appendix \ref{InstructionsForGrok} first for instructions on how to read this document and your role in its continuing composition.}

\chapter{Introduction}

This monograph seeks to provide philosophical underpinnings for the management of a large distributed shared repository of declarative knowledge.
Because of its philosophical concern with knowledge, the central impetus belongs to epistemology.
The approach in constructive, insofar as we are advocating the adoption of an abstract language for the representation of knowledge which has been devised to support formal deductive reasoning about the knowledge thus represented, and for that reason I call this an \emph{epistemological synthesis} (though that term embraces not just the notation, which is not original, but the supporting philosophical ideas.

That the approach is foundational means that the whole is constructed around a logical kernel, the exposition and support for which may be thought of as a philosophical kernel, which itself supports a broader philosophical outlook.
That epistemology is seen here is foundational, does not mean that it  alone provides the philosophical kernel, for we cannot speak of declarative knowledge without enlisting elements of metaphysical ontology, philosophical logic and the philosophy of language.
The logical kernel incorporates those philosophical elements into the syntax and semantics of an abstract declarative language which we believe to have universal properties desirable in a general and complete repository of declarative knowledge.

The philosophical support for such foundations would ideally address a variety of sceptical arguments which have been advanced against similar philosophical ideas.
Among these, addressing scepticism about semantics has a particular importance in this exposition.

It is my aim in this introduction to provide minimal historical context (which will be expanded later) for a first sketch of the core elements of the foundations, and the way in which these foundations support the superstructure.





Here's a sketch of what this introduction is going to cover.

\section{Meaning and Truth}

We stand, as I write, on the brink of achieving ``artificial intelligence'', with conversational intelligent systems exhibiting superhuman breadth of knowledge and command of language, but even less than human ability to separate truth from falsehood.
The limitations stem in part from the core objectives, essentially to replicate and then surpass human intelligence, achieving this by indiscriminate consumption of very large volumes of data from diverse sources.

Insofar as efforts have been made to impose standards on how these AIs work, it has been directed at `safety', without there being any clear conception of what that means, beyond avoiding catastrophic outomes.
Without much sign of progress on any reasonable metric of safety, the effect seems primarily to be one of alignment with the political leanings of their creators.

One company has prioritised `truth' rather than safety, but it is doubtful how they can achieve that objective when the human discourse on which AIs are trained exhibits a diversity of opinion, and fluidity of language which defies determination of truth or realisation of consistency.
At the same time, deception augmented by equivocation (motte and bailey) and large scale transformation of language sometimes designed to prevent the expression of dissenting view, have created a crisis of truth in which large political groupings have radically  contradictory perceptions of reality.
One might imagine in such circumstances that the design on cognitive artifacts would have as a priority the aim of establishing some solid baselines on truth, but even the one company which professes to care does not expose any insight into how that can be achieved.

\section{Epistemology}
  
Very early in the branch of philosophy which is now known as \emph{epistemology}, the theory of knowledge, Plato characterised knowledge as ``justified true belief''.
Can we look to epistemology for insights into how AIs which are truthful might be achieved?
This monograph is intended to contribute philosophical ideas about knowledge which might possibly enable AIs to escape or moderate the human susceptibility to equivocation, falsity and incoherence.

The transformation of purpose when we seek from epistemology philosophical underpinnings to enable rigour in superhuman synthetic intelligence is so great that the methods are likely to be unfamiliar and to represent a distinctive approach to epistemology.
I use ``synthetic epistemology'' to refer to the construction of epistemological theories of knowledge engineered to underpin some more practical purpose, and thence to title this monograph ``An Epistemological Synthesis''.

\section{A Short History of Deduction}

To expose the rational basis for the proposal which will eventually be detailed, it will be useful to provide some historical background.
First, in relation to the significance of deduction to the representation of knowledge, we note two major phases in the cultural evolution which has brought us to our present technological competence.
An ability to perform elementary deductive reasoning integral to an understanding of language, since that involves knowledge of basic conceptual inclusions, such as that horses are animals, which justifies the inference from there is a horse in the field to there is an animal in the field.

Elaborate chains of deductive reasoning are something else, and we only see this happening explicitly in the last three thousand years since the ancient Greeks abstracted from the arithmetic and geometric techniques of their predecessors into axiomatic mathematical theories.
This established a domain and a method which supported extended deductive reasoning for the first time, but outside those narrow confines deductive reasoning proved much less conclusive, giving philosophers motivation to extend the domain of effectiveness into their more abstruse subject matters.

Over the next three thousand years two important milestones underlined the strength of that motivation.
The first was Aristotle's Organon, the works on logic which articulated his conception of \emph{demonstrative science}, the second was the birth of modern science,

\section{Berlin's Three Legs}

In his essays on the origins of Romanticism \cite{berlinRR}, Isaiah Berlin, spoke of the Western Tradition as it stood before The Enlightenment, describing it as resting on three legs.
A part of that perception essentially articulated necessary conditions for a progressive accumulation of declarative knowledge 



\section{Sketch of Remainder of Introduction}
  
Here's the progression I anticipate:
\begin{itemize}
\item Say a bit more about demonstrative science, provide an account of demonstrative science and modern idea of hypothetico-deductive or nomologico-deductive science, showing the progression from the one to the other, and noting the impotence of Aristotle's syllogismic account of deduction and the power of numerical scientific theories.
\item Talk about Isaiah Berlin's account of the three legs upon which the western tradition rests, connecting this to the necessary conditions for the accumulation of large bodies of deductively applicable knowledge.
\item Note the doubtful position of empirical and evaluative knowledge in relation to those ideals, and the relevance of this to the shift from ``enlightenment'' to ``romanticism''.
\item Talk about the mitigation of reservations about the logical status of empirical science through the interpretation of science as providing abstract models rather than factual generalisations, and nod to similar treatment of evaluative language and practice.
\item Sketch the ongoing march of mathematicised science (despite romanticism) and the crisis in confidence in the rigour of mathmatical analysis leading to arithmetisation of analysis and the Frege's logicist foundations for mathematics.
  Talk about the role of logical kernels in that process and the dilemma between the idea of a single foundation, the apparent need for linguistic pluralism (partly driven by Carnap's attempt to extend the foundational method beyond mathematics into science).
\item Say something about the evolution of those mathematical foundations through Russell and Church to Gordon's HOL and its application to hardware verification.

\item Talk about the abstraction from Gordon's HOL which is the epicenter of the synthesis.
\item Enumerate the main claims supporting the relevance of this abstract foundation system to the development of appropriate knowedge repositories.
\item Enumerate the philosophical problems involved in justifying the claims and indicate the manner in which those problems will be addressed.
\item Sketch the how rest of the monograph chapter by chapter addresses those problems.
\end{itemize}

\section{The Structure of this Monograph}

Here a sketch of what is to follow in later chapters.


\cite{couturat1901logique}


\resetnotes

\ignore{
\chapter{A Philosophical Kernel}\label{PK}

%\input{kernel5.tex}
\resetnotes

\chapter{Evolution}

\cite{donald1991,murray2017evolution}

\chapter{Synthetic Epistemology and Foundational Abstract Semantics}

\input{epistem01.tex}
\resetnotes

\chapter{The Evolution of Declarative Knowledge}\label{EDK}

%\input{edk02.tex}
\resetnotes


\chapter{The Fundamental Triple Trichotomy}\label{FTT}

%\input{ftt01.tex}
\resetnotes
}%ignore

\appendix

\ignore{
\chapter{The Philosophy of Rudolf Carnap}\label{PRC}

\input{carnap01.tex}
\resetnotes

}%ignore

\appendix

\chapter{Instructions for Grok3}\label{InstructionsForGrok}

I have written this monograph with copious assistance from Grok 3, an AI chatbot from xAI.

Notwithstanding the invaluable contribution Grok has made, the writing is all mine.

The following section contains the general instructions which I have used to brief Grok.
Naturally, these instructions have evolved over the course of the writing and this their state at the end.
Different stages in the process demand different instructions, and I have tried to keep instructions of a general character  in a separate file (presented below), with supplementary files giving more specific asks for particular stages.

\section{General Guidance for Grok}

\input{grok3in.tex}

\listoftables
%\listoffigures

%\phantomsection
\addcontentsline{toc}{section}{Bibliography}
\bibliographystyle{rbjfmu}
\bibliography{rbj3}

%\phantomsection
\renewcommand{\indexname}{Index of Defined Terms}
\addcontentsline{toc}{section}{Index of Defined Terms}
{\twocolumn[]
{\small\printindex}}

%\vfill

%\tiny{
%Started 2024-10-19
%}%tiny


\end{document}

% LocalWords:
